master thesis

### 2018-03-02 Paper - Latex/Paper - Skizze - v2.tex
Paper - new title; add some references; Cross Evaluation

    	\vspace{20mm}
    	\begin{Large}
    	    \textbf{An analysis of the Castor package to predict parasitism in Eukaryota}\\
      \end{Large}

      \vspace{20mm}
      or

      \vspace{20mm}
    	\begin{Large}
    	    \textbf{Origins and losses of parasitism}\\
          an analysis of the phylogenetic tree of life with a parsimony-like algorithm\\
    	\end{Large}

@@

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{OTL}
      For our project we looked for a large database for phylogenetic trees and also for a taxonomic 
        tree. Since we run our algorithm on the phylogenetic tree, and for the evaluation and other 
        properties the taxonomy provides us with much more information. \\
      OTL gives us both. A synthesis of phylogenetic trees (currently 819 trees) and a taxonomic tree. 
        OTL also includes the large phylogenetic database TreeBASE \cite{Hinchliff2015}. \\
      \todo{Das steht auf der Website nicht in dem Paper...} \\
      For phylogenetic data, there are at least five big data collections, namely: ITIS (Integrated 
      Taxonomic Information System), NCBI (National Center for Biotechnology Information [18]), WORMS 
      (World Register of Marine Species [14]), GBIF (Global Biodiversity Information Facility [17]), 
      OTT (OpenTreeOfLife-Taxonomy [4]).
      \todotext{Marius:} "Every dataset has it's own characteristics and downsides. ITIS is only a small 
        set of 100\% confirmed and named species. GBIF is not composed with the help of phylogeny, the 
        same is valid for the NCBI taxonomy. The WORMS taxonomy is a way too small dataset of mostly 
        marine species. \\
      We choosed the taxonomy from OpenTreeOfLife because it's including most of the known taxonomies 
        and got synthesised by preffering taxonomies that match with available phylogenetic data. At the 
        same time the team from OTL preferred a maximum number of species [4]. This is resulting in 
        somekind of hybrid between taxonomy and phylogeny." \\

@@

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{GloBI}
    \todotext{Marius:} "There aren't many big active interaction databases out there, most of them are 
      offline or outdated. For example: Interaction Web Database [35], Webs on the Web [36], Animal 
      Diversity Web [37] and ecoweb [38]. GloBI is including most of the known ones and is still growing 
      actively [3]. So the question which interaction database could be used was answered rather quickly." \\ \\

      For tagging our leave nodes, we use the GloBI (Global Biotic Interactions) database. This database 
        consists of entries of the form: species A (source) interacts with B (target). \\
      We appointed some interactions, where we know from the biological perspective that the species 
      source or target has to be a parasite or a free-living species. These are the following:
      \begin{itemize}
        \item free-living source: preysOn, eats, flowersVisitedBy, hasPathogen, pollinatedBy, 
          hasParasite, hostOf
        \item free-living target: preyedUponBy, parasiteOf, visitsFlowersOf, pathogenOf, hasHost
        \item parasite source: parasiteOf, pathogenOf
        \item parasite target: hasParasite, hasPathogen
      \end{itemize}
      \todo{Interactions nochmal prüfen! Darauf basieren unsere Ergebnisse!} \\
      We build two lists: parasites and free-livings, and add the source or targets of an interaction
        to these. \\
      \todo{klar? Oder Beispiel bringen? (Katze isst Maus $\rightarrow$ Katze ist Freilebend)} \\
      % For every interaction from these lists, we get one or two species for our parasite or 
      %   free-living lists.
      \todo{einige spezies nicht mit einbezogen, da sie keine OTT id haben, hier könnte man noch 
        verbessern (future work)}
      \todo{You can find all interaction types here:
        https://github.com/jhpoelen/eol-globi-data/blob/master/eol-globi-lib/src/main/java/org/eol/globi/domain/InteractType.java .} \\ \\

@@

        \begin{center}
          \begin{longtable}{ |l|r|r| }
            \hline
            Model & Residuals Dev & Deviance Pr \\
            \hline \hline
            multifurc $\sim$ kingdom & 7774454 & \\
            multifurc $\sim$ phylum & 7435700 & 338754 \\
            multifurc $\sim$ class & 7337241 & 98459 \\
            \rowcolor{green} multifurc $\sim$ order & 7076068 & 261172 \\
            \hline \hline \hline
            multifurc $\sim$ phylum + depth & 7431609 & \\
            multifurc $\sim$ phylum + max.height & 7375889 & 55721 \\
            multifurc $\sim$ phylum + min.height & 7233486 & 142403 \\
            \rowcolor{green} multifurc $\sim$ phylum + mean.height & 7128318 & 105167 \\
            \hline \hline
            multifurc ~ phylum * depth & 7335396 & \\
            multifurc ~ phylum * max.height & 7311241 & 24155 \\
            multifurc ~ phylum * min.height & 7177002 & 134238 \\
            \rowcolor{green} multifurc ~ phylum * mean.height & 7020258 & 156745 \\
            \hline \hline
            multifurc ~ phylum & 7435700 & \\
            multifurc ~ phylum + mean.height & 7128318 & 307381 \\
            \rowcolor{green} multifurc ~ phylum * mean.height & 7020258 & 108061 \\
            \hline \hline \hline
            multifurc $\sim$ class + depth & 7334754 & \\
            multifurc $\sim$ class + max.height & 7275856 & 58898 \\
            multifurc $\sim$ class + min.height & 7144686 & 131170 \\
            \rowcolor{green} multifurc $\sim$ class + mean.height & 7055313 & 89374 \\
            \hline \hline
            multifurc ~ class * depth & 7250759 & \\
            multifurc ~ class * max.height & 7187504 & 63256 \\
            multifurc ~ class * min.height & 7094933 & 92570 \\
            \rowcolor{green} multifurc ~ class * mean.height & 6965794 & 129139 \\
            \hline \hline
            multifurc ~ class & 7337241 & \\
            multifurc ~ class + mean.height & 7055313 & 281928 \\
            \rowcolor{green} multifurc ~ class * mean.height & 6965794 & 89518 \\
            \hline \hline \hline
            multifurc $\sim$ order + depth & 7027578 & \\
            multifurc $\sim$ order + max.height & 7005424 & 22154 \\
            multifurc $\sim$ order + min.height & 6890703 & 114721 \\
            \rowcolor{green} multifurc $\sim$ order + mean.height & 6815271 & 75432 \\
            \hline
          \end{longtable}
        \end{center}
        * Residuals: Fehler - wieviele Werte sind nicht gut modelliert. (umso kleiner umso besser - grün) \\
        * Deviance - Abweichung \\

@@

@@

%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
\bibliography{bibliographie}
\bibliographystyle{alphadin}

>>>>

    	\vspace{20mm}
    	\begin{Large}
          \textbf{An analysis of maximum parsimony algorithms to predict parasitism in Eukaryota}\\
          using a large multiurcated phylogenetic synthesis tree
      \end{Large}

@@

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{OTL}
      For our project we looked for a large database for phylogenetic trees and also for a taxonomic 
        tree. Since we run our algorithm on the phylogenetic tree, and for the evaluation and other 
        properties the taxonomy provides us with much more information. \\
      OTL gives us both. A synthesis of phylogenetic trees (currently 819 trees) and a taxonomic tree. 
        OTL also includes the large phylogenetic database TreeBASE \cite{Hinchliff2015}. \\
      \todo{Das steht auf der Website nicht in dem Paper...} \\
      For phylogenetic data, there are at least five big data collections, namely: ITIS (Integrated 
        Taxonomic Information System) \cite{ITIS}, NCBI (National Center for Biotechnology Information) 
        \cite{NCBI1988}, WORMS (World Register of Marine Species) \cite{WoRMS2018}, GBIF (Global 
        Biodiversity Information Facility) \cite{GBIF}, OTT (OpenTreeOfLife-Taxonomy) 
        \cite{Hinchliff2015}. \\
      \todotext{Marius:} "Every dataset has it's own characteristics and downsides. ITIS is only a small 
        set of 100\% confirmed and named species. GBIF is not composed with the help of phylogeny, the 
        same is valid for the NCBI taxonomy. The WORMS taxonomy is a way too small dataset of mostly 
        marine species. \\
      We choosed the taxonomy from OpenTreeOfLife because it's including most of the known taxonomies 
        and got synthesised by preffering taxonomies that match with available phylogenetic data. At the 
        same time the team from OTL preferred a maximum number of species \cite{Hinchliff2015}. This is 
        resulting in somekind of hybrid between taxonomy and phylogeny." \\

@@

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{GloBI}
      \todotext{Marius:} "There aren't many big active interaction databases out there, most of them are 
        offline or outdated. For example: IWDB (Interaction Web Database) \cite{IWDB2003}, Webs on the 
        Web \cite{WOW2004}, Animal Diversity Web \cite{Myers2003} and ecoweb \cite{Cohen2010}. GloBI is 
        including most of the known ones and is still growing actively \cite{Poelen2014}. So the 
        question which interaction database could be used was answered rather quickly." \\

      This database consists of entries of the form: species A (source) interacts with B (target). \\
      We appointed some interactions, where we know from the biological perspective that the species 
      source or target has to be a parasite or a free-living species. These are the following:
      \begin{itemize}
        \item free-living source: preysOn, eats, flowersVisitedBy, hasPathogen, pollinatedBy, 
          hasParasite, hostOf
        \item free-living target: preyedUponBy, parasiteOf, visitsFlowersOf, pathogenOf, hasHost
        \item parasite source: parasiteOf, pathogenOf
        \item parasite target: hasParasite, hasPathogen
      \end{itemize}
      \todo{Interactions nochmal prüfen! Darauf basieren unsere Ergebnisse!} \\
      We build two lists: parasites and free-livings, and add the source or targets of an interaction
        to these. \\
      \todo{klar? Oder Beispiel bringen? (Katze isst Maus $\rightarrow$ Katze ist Freilebend)} \\
      % For every interaction from these lists, we get one or two species for our parasite or 
      %   free-living lists.
      \todo{einige spezies nicht mit einbezogen, da sie keine OTT id haben, hier könnte man noch 
        verbessern (future work)}
      \todo{You can find all interaction types here:
        https://github.com/jhpoelen/eol-globi-data/blob/master/eol-globi-lib/src/main/java/org/eol/globi/domain/InteractType.java .} \\ \\

@@

        \begin{center}
          \begin{longtable}{ |l|r|r| }
            \hline
            Model & Residuals Dev & Deviance Pr \\
            \hline \hline
            multifurc $\sim$ kingdom & 7774454 & \\
            multifurc $\sim$ phylum & 7435700 & 338754 \\
            multifurc $\sim$ class & 7337241 & 98459 \\
            \rowcolor{green!50} multifurc $\sim$ order & 7076068 & 261172 \\
            \hline \hline \hline
            multifurc $\sim$ phylum + depth & 7431609 & \\
            multifurc $\sim$ phylum + max.height & 7375889 & 55721 \\
            multifurc $\sim$ phylum + min.height & 7233486 & 142403 \\
            \rowcolor{green!50} multifurc $\sim$ phylum + mean.height & 7128318 & 105167 \\
            \hline \hline
            multifurc ~ phylum * depth & 7335396 & \\
            multifurc ~ phylum * max.height & 7311241 & 24155 \\
            multifurc ~ phylum * min.height & 7177002 & 134238 \\
            \rowcolor{green!50} multifurc ~ phylum * mean.height & 7020258 & 156745 \\
            \hline \hline
            multifurc ~ phylum & 7435700 & \\
            multifurc ~ phylum + mean.height & 7128318 & 307381 \\
            \rowcolor{green!50} multifurc ~ phylum * mean.height & 7020258 & 108061 \\
            \hline \hline \hline
            multifurc $\sim$ class + depth & 7334754 & \\
            multifurc $\sim$ class + max.height & 7275856 & 58898 \\
            multifurc $\sim$ class + min.height & 7144686 & 131170 \\
            \rowcolor{green!50} multifurc $\sim$ class + mean.height & 7055313 & 89374 \\
            \hline \hline
            multifurc ~ class * depth & 7250759 & \\
            multifurc ~ class * max.height & 7187504 & 63256 \\
            multifurc ~ class * min.height & 7094933 & 92570 \\
            \rowcolor{green!50} multifurc ~ class * mean.height & 6965794 & 129139 \\
            \hline \hline
            multifurc ~ class & 7337241 & \\
            multifurc ~ class + mean.height & 7055313 & 281928 \\
            \rowcolor{green!50} multifurc ~ class * mean.height & 6965794 & 89518 \\
            \hline \hline \hline
            multifurc $\sim$ order + depth & 7027578 & \\
            multifurc $\sim$ order + max.height & 7005424 & 22154 \\
            multifurc $\sim$ order + min.height & 6890703 & 114721 \\
            \rowcolor{green!50} multifurc $\sim$ order + mean.height & 6815271 & 75432 \\
            \hline
          \end{longtable}
        \end{center}
        * Residuals: Fehler - wieviele Werte sind nicht gut modelliert. (umso kleiner umso besser - grün) \\
        * Deviance - Abweichung \\

@@

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Cross evaluation - leave 100 out}
      We ran the castor algorithm (\todo{10000}) 10 times with leaving 100 randomized free-living or 
        parasitic species out of the input data to see how stable our result is. \\
      What could happen by removing a parasite or free-living of the list?
      \begin{itemize}
        \item It could be a specie, which don't exist in the tree leaf nodes. -> no effect
        \item It could be a specie, which exists in both lists. -> If it was a parasite, it is now 
          free-living, because we prefer parasites. Otherwise we have no effect again. (1053 are possible)
        \item Normal case: We loose information, because its a specie in our tree and we change it to a 
          leave node with no information.
      \end{itemize}
      \begin{table}
        \begin{center}
          \begin{tabular}{ |c||r|r|r||c|r| }
            \hline
            run & \multicolumn{3}{c||}{distance} & changed & lost \\
            & all nodes & leaf nodes & internal nodes & tag & tag \\
            \hline \hline
            0 & 4.0 & 4.0 & 0.0 & 4 & 48 \\
            1 & 4.0 & 4.0 & 0.0 & 2 & \cellcolor{green!50}45 \\
            2 & 85.3 & 79.3 & 6.0 & 3 & 51 \\
            3 & 1.0 & 1.0 & 0.0 & 1 & 47 \\
            \rowcolor{red!50}4 & 1248.0 & 1230.5 & 17.5 & \cellcolor{green!50}0 & 54 \\
            5 & 169.0 & 168.0 & 1.0 & 3 & 52 \\
            6 & 1.0 & 1.0 & 0.0 & 1 & 48 \\
            7 & 1.0 & 1.0 & 0.0 & 1 & 53 \\
            8 & 1.0 & 1.0 & 0.0 & 1 & \cellcolor{red!50}54 \\
            9 & 1.0 & 1.0 & 0.0 & 1 & 52 \\
            \hline
          \end{tabular}
        \end{center}
        \caption{Example runs. \\ green: min value, red: max value}
      \end{table}

      \begin{table}
        \begin{center}
          \begin{tabular}{ |cl||r|r|r|r| }
            \hline
            & & min & max & mean & variance \\
            \hline \hline
            \multirow{3}{*}{distance} & all & 1.0 & 1248.0 & 151.53 & 151528.92 \\
            & leaf nodes                    & 1.0 & 1230.5 & 149.08 & 147380.74 \\
            & internal nodes                & 0.0 & 17.5 & 2.45 & 31.47 \\ \hline
            \multicolumn{2}{|c||}{changed tag} & 0.0 & 4.0 & 1.7 & 1.57 \\ \hline
            lost tag & all tags             & 45.0 & 54.0 & \cellcolor{green!50}50.40 & 10.04 \\
            & FL tags                       & 16.0 & 32.0 & 23.10 & 18.77 \\
            & P tags                        & 16.0 & 36.0 & 27.30 & 41.34 \\
            \hline
          \end{tabular}
        \end{center}
        \caption{Statistics to Cross validation}
      \end{table}

@@

%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
\bibliography{bibliographie}
